Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-10-22 doc/Paper/main.tex
depth explained

      \item {\bf Survey depth}: we estimate the $10\sigma$ survey depth in different directions using a method similar to that outlined in \cite{2018PASJ...70S..25M}. In a given sky pixel, we collect all galaxies with $i$-band {\tt cmodel} fluxes measured with a signal-to-noise ratio between 9 and 11 (i.e. $9\le{\tt icmodel\_flux}/{\tt icmodel\_flux\_err}\le11$), and compute the mean of their $i$-band  {\tt cmodel} magnitude. We verified that we obtain similar depth estimates using other methods (e.g. magnitude corresponding to $10$ times the mean flux error), and that the magnitude limit of our sample ($i<24.5$) is bright enough that little or no area is lost in any of the fields by imposing this cut.

@@

>>>>

      \item {\bf Survey depth}: we estimate the $10\sigma$ survey depth in different directions as follows. For each pixel, we compute the average $i$-band flux error of all objects in it that are classified as stars. The corresponding survey depth is then given by this mean flux error multiplied by 10 and translated into a magnitude. Not all pixels contain enough stars to estimate a reliable mean flux error. We use a nearest-neighbor interpolation to assign a depth value to those pixels with fewer than 4 stars. We verified that we obtain similar depth estimates using other methods, such as that outlined in \cite{2018PASJ...70S..25M}, in which 10$\sigma$ depth is defined as the mean {\tt cmodel} magnitude of all sources with $i$-band signal-to-noise ratio between 9 and 11. It is worth emphasizing that our method uses only stars, discarding all sources classified as galaxies. The use of galaxies to estimate any systematics map is dangerous in the context of template deprojection, since the stochastic noise in the map associated with the discrete sampling of these sources correlates directly with the galaxy distribution, and can give rise to spurious correlations that bias the estimated power spectrum after deprojection (see Section \ref{ssec:methods.cell}). We provide further details in Appendix \ref{app:depth}.

@@

\appendix
\section{Reconstructing depth from discrete sources}\label{app:depth}
  The depth map used in our analysis (see Section \ref{ssec:methods.syst}) is constructed from the signal-to-noise ratios of sources classified as stars. The use of all sources, including galaxies, to generate this map, would result on spurious correlations with the true galaxy distribution that wbias the power spectra estimated from them. ould cause a bias in the power spectra after deprojecting it. We explore this effect further here.

  In a simplified model, let us assume that galaxies can be separated into discrete types labelled by an index $\alpha$. If we want to compute the value of a given quantity $Q$ in a pixel $p$ by averaging over all galaxies in that pixel, the result is:
  \begin{equation}
    \bar{Q}_p = \frac{\sum_\alpha N_{\alpha,p}Q_{\alpha,p}}{\sum_\alpha N_{\alpha,p}},
  \end{equation}
  where $N_{\alpha,p}$ is the number of galaxies of type $\alpha$ in pixel $p$, and $Q_{\alpha,p}$ is the value of $Q$ for those galaxies. Expanding $N_{\alpha,p}=\bar{N}_\alpha(1+\delta_{\alpha,p})$, where $\bar{N}_\alpha$ is the average number of galaxies of type $\alpha$ per pixel, and $\delta_{\alpha,p}$ is their overdensity, we obtain:
  \begin{align}\nonumber
    \bar{Q}_p&=\frac{\langle Q\rangle_p+\langle Q\delta\rangle_p}{1+\langle \delta\rangle_p}\\
             &\simeq\langle Q\rangle_p+\langle Q\delta\rangle_p-\langle Q\rangle_p\langle\delta\rangle_p+\mathcal{O}(\delta^2),
  \end{align}
  where we have defined $\langle f\rangle_p\equiv \sum_\alpha \bar{N}_\alpha f_{\alpha,p}/\sum_\alpha \bar{N}_\alpha$, and we have expanded to first order in $\delta$ in the second line. If $Q$ is constant across all galaxies, the second and third term in the second line cancel out, and the resulting map would receive no noise from the inhomogenous galaxy distribution ($\bar{Q}_p=\langle Q\rangle_p$) up to second order. This is however not the case for the flux errors used to estimate the depth map. Brighter sources have lower number densities (and therefore stronger shot-noise fluctuations), and also larger flux errors. Therefore, althought the depth map that results from using this method on galaxies is dominated by the large-scale depth fluctuations, the small-scale fluctuations trace the spatial distribution of the brightest sample, and would correlate with the true galaxy distribution. This violates the core assumption of template deprojection (that the deprojected contaminant templates do not correlate with the true map fluctuations), and cause a bias in the power spectrum.

  Our depth map is computed from star-classified objects and therefore, although it will be subject to the same small-scale noise due to this discrete sampling, these noise fluctuations do not correlate with the galaxy distribution, and therefore do not bias our results. A second-order correlation could arise from the small fraction of galaxies mis-classified as stars, but the associated effect is negligible given our uncertainties. Due to these complications, in the future it will be desirable to estimate the survey depth using image properties directly (e.g. by injecting fake galaxies, or using the background noise integrated over PSF-size apertures).
